Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2024-06-17 manuscript/manuscript.tex
addressed comments
minor polish

\abstract[Abstract]{The Kolmogorov--Smirnov (KS) test is a widely used 
statistical test that
assesses the conformity of a sample to a specified distribution. Its efficacy,
however, diminishes with serially dependent data and when parameters
within the hypothesized distribution are unknown. For independent data,
parametric and nonparametric bootstrap procedures are available to adjust for
estimated parameters. For serially dependent stationary data, parametric
bootstrap has been developed with a working serial dependence structure. A
counterpart for the nonparametric bootstrap approach, which needs a bias
correction, has not been studied. Addressing this gap, our study introduces a
bias correction method employing a nonparametric block bootstrap, which
approximates the distribution of the KS statistic accounting for unspecified
serial dependence and unspecified parameters. We assess its effectiveness
through simulations, scrutinizing both its size and power. The practicality of
our method is further illustrated with an examination of stock returns from the 
S\&P 500 index, showcasing its utility in real-world
applications.}

\keywords{bias-correction, p-value, time series, finance}

@@

%\eds{Please check consistency with how you refer to $F$.  In 
%Equation~\eqref{eq:ks_standard}, it is $F(x; \theta_0)$ where the data and 
%parameter arguments are separated by a semicolon.  At the beginning of 
%Section~\ref{sec:methods}, it is $F(\cdot \mid \theta)$, with the vertical bar
%separating the data argument from the parameter argument (and no 0 subscript
%either).   Then in this first 
%subsection it is $F(x; \hat\theta_n)$ with a semicolon separating the two 
%arguments.  Then immediately above its use is mixed with $F(x)$ without a 
%parameter argument (implying it is the true parameter $\theta$). This continues
%throughout the document. I suggest either using consistent notation throughout, 
%or explicitly defining any shortened/simplified notation at its first use.}
%\mc{switched everything to semicolons}
%\jy{replaced $F$ with $F(\cdot; \theta_0)$}

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters 
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - B_n(x) \Big\vert.
  \]
  where 
  $B_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is the known
  bias term.
\end{enumerate}

@@

\jy{Recover the definition of jy and eds for comments. Prepare two
  definitions in the source so that one can be easily commented out.}
\mc{addressed}

\jy{Cite Kunsch (1989) in the proof as appropriate}

\jy{In the Supp: make sure notations are consistent (e.g., use
  $\{X_t: t = 1, \ldots, n\}$ for observed data);
  clean up the styles (linewidth etc.)}
\mc{addressed}

@@

\begin{table}[tbp]
\centering
\caption{Empirical sizes (ratio of rejection rate to significance level)
for the test that sample follows its true distribution family for
different values of AR(1) coefficient and for different 
significance levels.
\eds{Not sure if it will help presentation or interpretation, but you could
  consider presenting the values inside the table as estimated size/true size}
\mc{addressed}
\eds{I meant reporting the ratio, e.g, dividing all values currently in the    first row by 0.01, dividing all values in the second row by 0.05. Sometimes this
	is helpful if there is a conservative or liberal pattern, which there seems
	to be as $|\tau|$ increases away from zero, but it may be fine to leave it as
	is.}\mc{how does this look?}}
\label{table:size}
\begin{tabular}{lllrrrrrrr}
  \toprule
$F(\cdot; \theta)$ & $n$ & $\alpha$ & $\tau = -0.75$ & $\tau = -0.5$ & $\tau = -0.25$ 
& $\tau = 0$ & $\tau = 0.25$ & $\tau = 0.5$ & $\tau = 0.75$ \\ 
  \midrule
{$N(8,8)$} & {100} & 0.01 & 5.210 & 1.240 & 1.120 & 1.060 & 1.140 & 1.580 & 5.630 \\ 
& & 0.05 & 2.940 & 1.256 & 1.090 & 1.096 & 1.150 & 1.450 & 3.410 \\ 
& & 0.10 & 2.277 & 1.186 & 1.147 & 1.072 & 1.154 & 1.424 & 2.714 \\ 
& {200} & 0.01 & 4.390 & 1.130 & 0.900 & 1.140 & 1.090 & 1.380 & 5.900 \\ 
& & 0.05 & 2.556 & 1.120 & 1.020 & 1.064 & 1.092 & 1.382 & 3.452 \\ 
& & 0.10 & 2.080 & 1.123 & 1.069 & 1.086 & 1.051 & 1.332 & 2.650 \\ 
& {400} & 0.01 & 3.240 & 1.070 & 0.990 & 1.080 & 0.970 & 1.260 & 4.670 \\ 
& & 0.05 & 2.032 & 1.000 & 0.994 & 1.120 & 0.978 & 1.184 & 2.836 \\ 
& & 0.10 & 1.734 & 1.023 & 1.002 & 1.089 & 1.017 & 1.192 & 2.232 \\ 
& {800} & 0.01 & 2.520 & 0.900 & 0.980 & 1.110 & 1.020 & 1.350 & 3.540 \\ 
& & 0.05 & 1.766 & 1.008 & 0.972 & 0.956 & 0.962 & 1.236 & 2.438 \\ 
& & 0.10 & 1.555 & 1.026 & 0.986 & 0.961 & 0.997 & 1.206 & 2.044 \\ [1ex]
{$\Gamma(8,1)$} & {100} & 0.01 & 4.880 & 1.420 & 1.170 & 1.240 & 1.110 & 1.610 & 5.760 \\ 
  & & 0.05 & 2.850 & 1.282 & 1.222 & 1.170 & 1.178 & 1.464 & 3.374 \\ 
  & & 0.10 & 2.241 & 1.224 & 1.147 & 1.175 & 1.150 & 1.375 & 2.679 \\ 
  & {200} & 0.01 & 4.080 & 1.000 & 0.890 & 1.220 & 1.160 & 1.430 & 6.080 \\ 
  & & 0.05 & 2.510 & 1.118 & 1.014 & 1.068 & 1.116 & 1.290 & 3.380 \\ 
  & & 0.10 & 1.991 & 1.105 & 1.022 & 1.048 & 1.072 & 1.233 & 2.664 \\ 
  & {400} & 0.01 & 3.140 & 0.970 & 1.140 & 0.940 & 1.030 & 1.310 & 4.300 \\ 
  & & 0.05 & 2.004 & 0.964 & 1.042 & 1.062 & 1.122 & 1.236 & 2.808 \\ 
  & & 0.10 & 1.697 & 1.025 & 1.066 & 1.018 & 1.100 & 1.177 & 2.232 \\ 
  & {800} & 0.01 & 2.490 & 0.920 & 1.130 & 1.120 & 1.000 & 1.130 & 3.810 \\ 
  & & 0.05 & 1.804 & 0.976 & 1.064 & 1.022 & 1.086 & 1.196 & 2.474 \\ 
  & & 0.10 & 1.561 & 1.024 & 1.072 & 1.034 & 1.067 & 1.177 & 1.997 \\ 
   \bottomrule
\end{tabular}
\end{table}

For a closer scrutiny on whether the test holds its size at commonly used
significance levels 0.01, 0.05, and 0.10, Table~\ref{table:size}
summarizes the empirical sizes of the tests based on 10000 replicates.
\eds{Is 1000 enough to truly evaluate whether size is good at the 0.01 level?}
The agreement between the empirical sizes and the nominal size
improves as sample size improves, as expected. For $\tau = -0.75$
and $\tau = 0.75$, a sample size as large as 800 is still not sufficient for the
agreement to be reached. For weaker dependence levels, our approach is working
for $-0.25 < \tau < 0.25$ for $n = 100$. However, a larger sample size seems 
necessary, greater than 400 perhaps, for $\tau$ as strong as $0.5$. To 
summarize, for weaker dependence, our method works without need for a 
particularly large sample size, whereas for moderate dependence, a larger sample
size is required. For practical situations with a $\tau$ of magnitude 0.5 or
lower, it is still reasonable to recommend our method with $n \ge 400$.
This recommendation echoes the high sample size requirement for
confidence intervals constructed with block bootstrapping in
estimating location, scale, and autocorrelation parameters of
stationary series~\citep{chandy2024sample}.

@@

\eds{I think it would be better to show in simulation first (perhaps 
in supplemental materials if we are concerned with paper length) that the other
methods are not valid or unreliable under this setting, and then 
further discredit those results here in the real data analysis.
(I would expect a reviewer to ask for this comparison anyway.)}
\jy{we could design scenarios where SPBB fails: some dependency
  structure that cannot be well-approximated by ARMA.}

@@

None reported.

>>>>

\abstract[Abstract]{The Kolmogorov--Smirnov (KS) test is a widely used 
statistical test that
assesses the conformity of a sample to a specified distribution. Its efficacy,
however, diminishes with serially dependent data and when parameters
within the hypothesized distribution are unknown. For independent data,
parametric and nonparametric bootstrap procedures are available to adjust for
estimated parameters. For serially dependent stationary data, parametric
bootstrap has been developed with a working serial dependence structure. A
counterpart for the nonparametric bootstrap approach, which needs a bias
correction, has not been studied. Addressing this gap, our study introduces a
bias correction method employing a nonparametric block bootstrap, which
approximates the distribution of the KS statistic in assessing the
goodness-of-fit of the marginal distribution of a stationary series,
accounting for unspecified
serial dependence and unspecified parameters. We assess its effectiveness
through simulations, scrutinizing both its size and power. The practicality of
our method is further illustrated with an examination of stock returns from the 
S\&P 500 index, showcasing its utility in real-world
applications.}

% alphabetical; no repeating anything in the title already
\keywords{bias-correction, financial data, marginal distribution,
  stationary series, time series}

@@

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters 
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - B_n(x) \Big\vert.
  \]
  where 
  $B_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is the estimated
  bias term.
\end{enumerate}

@@

@@

\begin{table}[tbp]
\centering
\caption{Empirical sizes (ratio of rejection rate to significance level)
for the test that sample follows its true distribution family for
different values of AR(1) coefficient and for different 
significance levels.
\eds{Not sure if it will help presentation or interpretation, but you could
  consider presenting the values inside the table as estimated size/true size}
\mc{addressed}
\eds{I meant reporting the ratio, e.g, dividing all values currently in the    first row by 0.01, dividing all values in the second row by 0.05. Sometimes this
	is helpful if there is a conservative or liberal pattern, which there seems
	to be as $|\tau|$ increases away from zero, but it may be fine to leave it as
	is.}\mc{how does this look?}
\jy{Once finalized, don't forget to update the caption.}
    }
\label{table:size}
\begin{tabular}{lllrrrrrrr}
  \toprule
$F(\cdot; \theta)$ & $n$ & $\alpha$ & $\tau = -0.75$ & $\tau = -0.5$ & $\tau = -0.25$ 
& $\tau = 0$ & $\tau = 0.25$ & $\tau = 0.5$ & $\tau = 0.75$ \\ 
  \midrule
{$N(8,8)$} & {100} & 0.01 & 5.210 & 1.240 & 1.120 & 1.060 & 1.140 & 1.580 & 5.630 \\ 
& & 0.05 & 2.940 & 1.256 & 1.090 & 1.096 & 1.150 & 1.450 & 3.410 \\ 
& & 0.10 & 2.277 & 1.186 & 1.147 & 1.072 & 1.154 & 1.424 & 2.714 \\ 
& {200} & 0.01 & 4.390 & 1.130 & 0.900 & 1.140 & 1.090 & 1.380 & 5.900 \\ 
& & 0.05 & 2.556 & 1.120 & 1.020 & 1.064 & 1.092 & 1.382 & 3.452 \\ 
& & 0.10 & 2.080 & 1.123 & 1.069 & 1.086 & 1.051 & 1.332 & 2.650 \\ 
& {400} & 0.01 & 3.240 & 1.070 & 0.990 & 1.080 & 0.970 & 1.260 & 4.670 \\ 
& & 0.05 & 2.032 & 1.000 & 0.994 & 1.120 & 0.978 & 1.184 & 2.836 \\ 
& & 0.10 & 1.734 & 1.023 & 1.002 & 1.089 & 1.017 & 1.192 & 2.232 \\ 
& {800} & 0.01 & 2.520 & 0.900 & 0.980 & 1.110 & 1.020 & 1.350 & 3.540 \\ 
& & 0.05 & 1.766 & 1.008 & 0.972 & 0.956 & 0.962 & 1.236 & 2.438 \\ 
& & 0.10 & 1.555 & 1.026 & 0.986 & 0.961 & 0.997 & 1.206 & 2.044 \\ [1ex]
{$\Gamma(8,1)$} & {100} & 0.01 & 4.880 & 1.420 & 1.170 & 1.240 & 1.110 & 1.610 & 5.760 \\ 
  & & 0.05 & 2.850 & 1.282 & 1.222 & 1.170 & 1.178 & 1.464 & 3.374 \\ 
  & & 0.10 & 2.241 & 1.224 & 1.147 & 1.175 & 1.150 & 1.375 & 2.679 \\ 
  & {200} & 0.01 & 4.080 & 1.000 & 0.890 & 1.220 & 1.160 & 1.430 & 6.080 \\ 
  & & 0.05 & 2.510 & 1.118 & 1.014 & 1.068 & 1.116 & 1.290 & 3.380 \\ 
  & & 0.10 & 1.991 & 1.105 & 1.022 & 1.048 & 1.072 & 1.233 & 2.664 \\ 
  & {400} & 0.01 & 3.140 & 0.970 & 1.140 & 0.940 & 1.030 & 1.310 & 4.300 \\ 
  & & 0.05 & 2.004 & 0.964 & 1.042 & 1.062 & 1.122 & 1.236 & 2.808 \\ 
  & & 0.10 & 1.697 & 1.025 & 1.066 & 1.018 & 1.100 & 1.177 & 2.232 \\ 
  & {800} & 0.01 & 2.490 & 0.920 & 1.130 & 1.120 & 1.000 & 1.130 & 3.810 \\ 
  & & 0.05 & 1.804 & 0.976 & 1.064 & 1.022 & 1.086 & 1.196 & 2.474 \\ 
  & & 0.10 & 1.561 & 1.024 & 1.072 & 1.034 & 1.067 & 1.177 & 1.997 \\ 
   \bottomrule
\end{tabular}
\end{table}

For a closer scrutiny on whether the test holds its size at commonly used
significance levels 0.01, 0.05, and 0.10, Table~\ref{table:size}
summarizes the empirical sizes of the tests based on 10000 replicates.
The agreement between the empirical sizes and the nominal size
improves as sample size improves, as expected. For $\tau = -0.75$
and $\tau = 0.75$, a sample size as large as 800 is still not sufficient for the
agreement to be reached. For weaker dependence levels, our approach is working
for $-0.25 < \tau < 0.25$ for $n = 100$. However, a larger sample size seems 
necessary, greater than 400 perhaps, for $\tau$ as strong as $0.5$. To 
summarize, for weaker dependence, our method works without need for a 
particularly large sample size, whereas for moderate dependence, a larger sample
size is required. For practical situations with a $\tau$ of magnitude 0.5 or
lower, it is still reasonable to recommend our method with $n \ge 400$.
This recommendation echoes the high sample size requirement for
confidence intervals constructed with block bootstrapping in
estimating location, scale, and autocorrelation parameters of
stationary series~\citep{chandy2024sample}.

@@

% \eds{I think it would be better to show in simulation first (perhaps 
% in supplemental materials if we are concerned with paper length) that the other
% methods are not valid or unreliable under this setting, and then 
% further discredit those results here in the real data analysis.
% (I would expect a reviewer to ask for this comparison anyway.)}
% \jy{We could design scenarios where SPBB fails: some dependency
%   structure that cannot be well-approximated by ARMA. Could do this
%   if requested.}

@@

JY and XZ's research was partially supported by NSF grant
DMS2210735.
